Types for R, Empirically Designed and Evaluated

### 2020-05-09 paper/main.tex
disparate labels, and comments
cleanup

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{my_style}

\usepackage{listings, wrapfig}

\lstset{ 
  language=R,                     % the language of the code
  basicstyle=\tiny\ttfamily, % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\tiny\color{blue},  % the style that is used for the line-numbers
  stepnumber=1,                   % the step between two line-numbers. If it is 1, each line
                                  % will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,                   % adds a frame around the code
  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  keywordstyle=\color{teal},      % keyword style
  commentstyle=\color{olive},   % comment style
  stringstyle=\color{magenta}      % string literal style
}

\newcommand{\code}[1]{\text{\lstinline[basicstyle=\ttfamily\smaller]~#1~}}

% Macros for type names from the old paper.
\newcommand{\attr}[2]{\ensuremath{#1_{\mathtt{#2}}}\xspace}
\newcommand{\attrclass}[3]{\ensuremath{#1^{\mathtt{#3}}_{\mathtt{#2}}}\xspace}
\renewcommand{\to}{\ensuremath{\rightarrow}\xspace}
\newcommand{\D}{\ensuremath{\small\vec{\mathtt D}}\xspace} % Double
\newcommand{\I}{\ensuremath{\small\vec{\mathtt I}}\xspace} % Integer
\renewcommand{\C}{\ensuremath{\small\vec{\mathtt C}}\xspace} % Character
\renewcommand{\L}{\ensuremath{\small\vec{\mathtt L}}\xspace} % Logical
\newcommand{\R}{\ensuremath{\small\vec{\mathtt R}}\xspace} % Raw
\newcommand{\X}{\ensuremath{\small\vec{\mathtt X}}\xspace} % Complex
\newcommand{\Y}{\ensuremath{\small\vec{\mathtt Y}}\xspace} % Symbol
\newcommand{\sY}{\ensuremath{\small{\mathtt Y}}\xspace} % Symbol
\newcommand{\sS}{\ensuremath{\small{\mathtt S}}\xspace} % S4
\newcommand{\sF}{\ensuremath{\small{\mathtt F}}\xspace} % Closure
\newcommand{\sE}{\ensuremath{\small{\mathtt E}}\xspace} % Env
\renewcommand{\R}{\ensuremath{\small\vec{\mathtt R}}\xspace} % Raw
\newcommand{\sN}{\ensuremath{\small{\mathtt N}}\xspace}     % Null
%\renewcommand{\l}{\ensuremath{\small L<?>}\xspace}     % List
\renewcommand{\l}{\ensuremath{\small\underline{\mathtt ?}}\xspace}     % List
\newcommand{\sD}{\ensuremath{\small{\mathtt D}}\xspace} % Double
\newcommand{\sI}{\ensuremath{\small{\mathtt I}}\xspace} % Integer
\newcommand{\sC}{\ensuremath{\small{\mathtt C}}\xspace} % Character
\newcommand{\sL}{\ensuremath{\small{\mathtt L}}\xspace} % Logical
\newcommand{\sX}{\ensuremath{\small{\mathtt X}}\xspace} % Complex
\newcommand{\sR}{\ensuremath{\small{\mathtt R}}\xspace} % Raw
\newcommand{\ANY}{\ensuremath{\small{\mathtt ?}}\xspace}     % Any
\newcommand{\lT}[1]{\ensuremath{\small\underline{\mathtt{#1}}}\xspace}     % list<T>
\newcommand{\M}[1]{\ensuremath{\attr{\vec{\tt #1}}{mat}}\xspace}     % matrix
\newcommand{\df}{\ensuremath{\attr{\l}{df}}\xspace}     % data.frame

@@

\begin{document}

%% Title information
\title[]{Types for R, Empirically Designed and Evaluated}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
% \titlenote{}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
% \subtitle{}                     %% \subtitle is optional
% \subtitlenote{}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}

\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

@@

\subsection{contractr}
\label{sec:contractr}
\contractr is an R package that implements our type system for the R
package ecosystem. It adds contracts to package functions for type-checking
arguments and results on function calls. On contract failure, it reports a type
mismatch warning with the package name, function name, argument name, argument
position, expected type signature, actual type signature of the value observed,
and a stack trace for the called function. \contractr works by modifying
function definition to insert a call to the type-checking function. When the
modified function is called, the type-checking function is invoked on the
modified function's arguments and return value.

@@

%
%
%
%
\subsection{The Corpus}

@@

In order to ensure that our type system design aligns with the day-to-day usage of R, we performed a corpus analysis of some of the most widely used R packages.
We built our dynamic analysis in the R-dyntrace~\cite{oopsla19} dynamic analysis framework, which is itself built on a modified version of R targeting R version 3.5.0---we refer the reader to Section~\ref{sec:r-dyntrace} for more details on R-dyntrace.

@@

%
%
%
%
\subsection{Threats to Validity}
\label{sec:threatstovalidity}

%
%
%
%
\subsection{Future Work}
\label{sec:futurework}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
\bibliography{bib/biblio,bib/jv,bib/r,bib/new,bib/gradual}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

>>>>

\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\setcopyright{none}
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{my_style}
\usepackage{listings, wrapfig,xspace}

\lstset{language=R}
\definecolor{LightGray}{rgb}{.92,.92,.92}
\definecolor{Gray}{rgb}{.3,.3,.3}
\definecolor{DarkGray}{rgb}{.5,.5,.5}
\lstset{ %
  columns=flexible,
  captionpos=b,
  frame=single,
  framerule=0pt,
  tabsize=2,
  belowskip=0.5em,
  backgroundcolor=\color{LightGray},
  basicstyle=\small\ttfamily,
  emphstyle=,
  keywordstyle=,
  commentstyle=\color{Gray}\em,
  stringstyle=\color{Gray}
}
\lstdefinestyle{R}{ %
  language=R,
  morekeywords={assign, delayedAssign},
  deletekeywords={env, equal, c, runif, trace, args, exp, t, all},
  breaklines=true
}
\lstdefinestyle{Rin}{ %
  style=R,
  breaklines=false
}

\newcommand{\eg}{\emph{e.g.},\xspace}
\newcommand{\ie}{\emph{i.e.},\xspace}

\newcommand{\PIR}{\textsf{PIR}\xspace}
\newcommand\pirI[1]{\mathtt{#1}}
\renewcommand{\c}[1]{{\lstinline[style=Rin]!#1!}\xspace}
\newcommand{\code}[1]{{\lstinline[style=Rin]!#1!}\xspace}
% Macros for type names from the old paper.
\newcommand{\attr}[2]{\ensuremath{#1_{\mathtt{#2}}}\xspace}
\newcommand{\attrclass}[3]{\ensuremath{#1^{\mathtt{#3}}_{\mathtt{#2}}}\xspace}
\renewcommand{\to}{\ensuremath{\rightarrow}\xspace}
\newcommand{\D}{\ensuremath{\small\vec{\mathtt D}}\xspace} % Double
\newcommand{\I}{\ensuremath{\small\vec{\mathtt I}}\xspace} % Integer
\renewcommand{\C}{\ensuremath{\small\vec{\mathtt C}}\xspace} % Character
\renewcommand{\L}{\ensuremath{\small\vec{\mathtt L}}\xspace} % Logical
\newcommand{\R}{\ensuremath{\small\vec{\mathtt R}}\xspace} % Raw
\newcommand{\X}{\ensuremath{\small\vec{\mathtt X}}\xspace} % Complex
\newcommand{\Y}{\ensuremath{\small\vec{\mathtt Y}}\xspace} % Symbol
\newcommand{\sY}{\ensuremath{\small{\mathtt Y}}\xspace} % Symbol
\newcommand{\sS}{\ensuremath{\small{\mathtt S}}\xspace} % S4
\newcommand{\sF}{\ensuremath{\small{\mathtt F}}\xspace} % Closure
\newcommand{\sE}{\ensuremath{\small{\mathtt E}}\xspace} % Env
\renewcommand{\R}{\ensuremath{\small\vec{\mathtt R}}\xspace} % Raw
\newcommand{\sN}{\ensuremath{\small{\mathtt N}}\xspace}     % Null
%\renewcommand{\l}{\ensuremath{\small L<?>}\xspace}     % List
\renewcommand{\l}{\ensuremath{\small\underline{\mathtt ?}}\xspace}     % List
\newcommand{\sD}{\ensuremath{\small{\mathtt D}}\xspace} % Double
\newcommand{\sI}{\ensuremath{\small{\mathtt I}}\xspace} % Integer
\newcommand{\sC}{\ensuremath{\small{\mathtt C}}\xspace} % Character
\newcommand{\sL}{\ensuremath{\small{\mathtt L}}\xspace} % Logical
\newcommand{\sX}{\ensuremath{\small{\mathtt X}}\xspace} % Complex
\newcommand{\sR}{\ensuremath{\small{\mathtt R}}\xspace} % Raw
\newcommand{\ANY}{\ensuremath{\small{\mathtt ?}}\xspace}     % Any
\newcommand{\lT}[1]{\ensuremath{\small\underline{\mathtt{#1}}}\xspace}     % list<T>
\newcommand{\M}[1]{\ensuremath{\attr{\vec{\tt #1}}{mat}}\xspace}     % matrix
\newcommand{\df}{\ensuremath{\attr{\l}{df}}\xspace}     % data.frame

@@

\begin{document}
\title{Types for R, Empirically Designed and Evaluated}
\begin{abstract}

\end{abstract}
\maketitle

@@

\subsection{contractr}
\label{sec:contractr}

\AT{I think we should move this to the Methodology section, which we should maybe rename to implementation. Thoughts?}

\contractr is an R package that implements our type system for the R
package ecosystem. It adds contracts to package functions for type-checking
arguments and results on function calls. On contract failure, it reports a type
mismatch warning with the package name, function name, argument name, argument
position, expected type signature, actual type signature of the value observed,
and a stack trace for the called function. \contractr works by modifying
function definition to insert a call to the type-checking function. When the
modified function is called, the type-checking function is invoked on the
modified function's arguments and return value.

@@

%
%
%
%
\subsection{The Corpus}
\label{sec:corpus}

@@

In order to ensure that our type system design aligns with the day-to-day usage of R, we performed a corpus analysis of some of the most widely used R packages.
The corpus is discussed in detail in Section~\ref{sec:corpus}.
We built our dynamic analysis in the R-dyntrace~\cite{oopsla19} dynamic analysis framework, which is itself built on a modified version of R targeting R version 3.5.0---we refer the reader to Section~\ref{sec:r-dyntrace} for more details on R-dyntrace.

@@

%
%
%
%
\subsection{Threats to Validity}

\subsection{Future Work}

\bibliography{bib/biblio,bib/jv,bib/r,bib/new,bib/gradual}

\end{document}
\appendix
\section{Appendix}
